---
title: "ITS_Project"
author: "Ishaan_Pahuja"
date: "2026-04-13"
output: pdf_document
---

```{r}
df <- read.csv("olympic.csv")
```

## SRS

```{r}
set.seed(999)
library(dplyr)

data <- read.csv("olympic.csv")

# 1. Simple Random Sample
# n=200

set.seed(999)
srs_sample <- data[sample(nrow(data), 200), ]

```

## Stratified

```{r}
# 2. Stratified Sample
# n=100 for female and n=100 for male

set.seed(999)
stratified_sample <- data %>%
  group_by(GENDER) %>%
  sample_n(size = 100) %>%
  ungroup()
```

## One Stage Cluster, SRS

```{r}
set.seed(999)

data$SPORT <- trimws(data$SPORT)

sports <- unique(data$SPORT)
cluster_sizes <- table(data$SPORT)

# SRS of 12 clusters, then all athletes in those clusters
selected_srs1      <- sample(sports, 12)
cluster_srs_sample <- data[data$SPORT %in% selected_srs1, ]
```

## One Stage Cluster, PPS

```{r}
set.seed(999)

# PPS of 12 clusters (prob proportional to size), then all athletes
selected_pps1 <- sample(names(cluster_sizes), 12,
                        prob = as.numeric(cluster_sizes), replace = FALSE)
cluster_pps_sample <- data[data$SPORT %in% selected_pps1, ]
```

## Stratified Clustering

```{r}
set.seed(999)

female_data     <- data[data$GENDER == "F", ]
sel_female_cl   <- sample(unique(female_data$SPORT), 12)
strat_cl_female <- female_data[female_data$SPORT %in% sel_female_cl, ]

male_data       <- data[data$GENDER == "M", ]
sel_male_cl     <- sample(unique(male_data$SPORT), 12)
strat_cl_male   <- male_data[male_data$SPORT %in% sel_male_cl, ]

strat_cluster_sample <- rbind(strat_cl_female, strat_cl_male)
```

## Two Stage Cluster, SRS

```{r}
set.seed(999)

get_mi <- function(Mi) {
  if      (Mi <= 5)  Mi - 1
  else if (Mi <= 20) floor(Mi / 2)
  else if (Mi <= 50) floor(Mi / 3)
  else               floor(Mi / 4)
}

selected_2srs <- sample(sports, 18)

two_stage_srs_sample <- do.call(rbind, lapply(selected_2srs, function(sp) {
  cl <- data[data$SPORT == sp, ]
  mi <- get_mi(nrow(cl))
  if (mi == 0) return(NULL)
  cl[sample(nrow(cl), mi), ]
}))

```
Description:
The sampling for Two Stage Clustering, with SRS within clusters is done via r. First I created a helper function to get the within cluster sample size via rules stated on the instruction docs. Then I sampled n=18 sports from a list of all the sports which represents the N clusters. After completing the first stage and obtaining the clusters, for each cluster I filtered the rows of the original data that have the sport/belong in the cluster, and called the helper function on the size of the cluster, Mi. Utilizing the mi, cluster sample size, I sampled the cluster via the included sample function.


Sample Statistics
```{r}
cluster_stats <- two_stage_srs_sample %>%
  group_by(SPORT) %>%
  summarise(
    mi = n(),
    ybar = mean(HEIGHT),
    s_sqr = var(HEIGHT),
    p_hat = mean(`OLYMPIC.EXPERIENCE` == "Yes")
  ) %>%
  left_join(
    data %>%
      group_by(SPORT) %>%
      summarise(Mi = n()), 
    by = "SPORT"
  ) %>%
  select(SPORT, Mi, mi, ybar, s_sqr, p_hat)
```

Parameter Estimates
```{r}
Mi    <- cluster_stats$Mi
mi    <- cluster_stats$mi
ybar  <- cluster_stats$ybar
s_sqr <- cluster_stats$s_sqr
p_hat <- cluster_stats$p_hat

N <- 30
n <- 18
M_bar_est <- mean(Mi)                  

mu_r_hat <- sum(ybar * Mi) / sum(Mi)
p_r_hat <- sum(p_hat * Mi) / sum(Mi)

sr2_mu <- sum((Mi * ybar - Mi * mu_r_hat)^2) / (n - 1)
sr2_p  <- sum((Mi * p_hat - Mi * p_r_hat)^2) / (n - 1)

term1_mu <- (1 - n/N) * (sr2_mu / (n * M_bar_est^2))
term2_mu <- (1 / (n * N * M_bar_est^2)) * sum(Mi^2 * (1 - mi/Mi) * (s_sqr / mi))
var_mu_r <- term1_mu + term2_mu

q_hat <- 1 - p_hat
term1_p <- (1 - n/N) * (sr2_p / (n * M_bar_est^2))
term2_p_elements <- ifelse(mi > 1, 
                           Mi^2 * (1 - mi/Mi) * ((p_hat * q_hat) / (mi - 1)), 
                           0)
term2_p <- (1 / (n * N * M_bar_est^2)) * sum(term2_p_elements)
var_p_r <- term1_p + term2_p

bound_mu_r <- 2 * sqrt(var_mu_r)
bound_p_r  <- 2 * sqrt(var_p_r)

results_ratio <- data.frame(
  Parameter = c("Mean Height", "Proportion"),
  Estimate = c(mu_r_hat, p_r_hat),
  Variance = c(var_mu_r, var_p_r),
  Bound_of_Error = c(bound_mu_r, bound_p_r)
)

```

```{r}
mean(data$HEIGHT)
mean(data$OLYMPIC.EXPERIENCE == "Yes")
```

#

## Two Stage Cluster, PPS

```{r}
set.seed(999)

selected_2pps <- sample(names(cluster_sizes), 18,
                        prob = as.numeric(cluster_sizes), replace = FALSE)

two_stage_pps_sample <- do.call(rbind, lapply(selected_2pps, function(sp) {
  cl <- data[data$SPORT == sp, ]
  mi <- get_mi(nrow(cl))
  if (mi == 0) return(NULL)
  cl[sample(nrow(cl), mi), ]
}))



```
